# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Read data in *.zip* format, and then convert it into a data table for program
efficiency.
```{r echo=TRUE}
library(data.table)
file <- unz("activity.zip", "activity.csv")
data <- data.table(read.table(file, header=TRUE, sep=",", na.string="NA"))
summary(data)
class(data)
```


## What is mean total number of steps taken per day?

Calculate the sum of steps group by date, and store the result into *stepsum*.

Notice that the *NA* values are removed in the *sum* function, so if the data is
missing on a specific date, it will result in a *zero sum value*.
```{r}
stepsum <- data[, sum(steps, na.rm=TRUE), by=date]
setnames(stepsum, c("date", "stepsum"))
```

We can get the histogram of the total number of steps taken each day as follows.
```{r}
hist(stepsum$stepsum, xlab="Total Number of Steps Taken Each Day", 
     main="Histogram of the Total Number of Steps Taken Each Day",
     col="blue")
```

**Mean**:
```{r}
mean(stepsum$stepsum, na.rm=TRUE)
```

**Median**:
```{r}
median(stepsum$stepsum, na.rm=TRUE)
```


## What is the average daily activity pattern?
The time series plot can be made by this:
```{r}
stepmean <- data[, mean(steps, na.rm=TRUE), by=interval]
setnames(stepmean, c("interval", "stepmean"))
plot(stepmean, type="l", ylab="Average Number of Steps Taken")
```
The interval that contains the maximum number of steps can be evaluated via:
```{r}
stepmean[sort(stepmean, decreasing=TRUE)][1, interval]
```


## Imputing missing values
The total number of missing values in the dataset is calculated by:
```{r}
data[, sum(is.na(steps))]
```
We are going to fill the NAs with the average steps taken within the interval,
which has been calculated and stored in *stepmean*.

First of all we will define a function for filling the NA steps.
```{r}
fillna <- function(steps, stepmean) {
   if (!is.na(steps)) {
       steps
   } else {
       stepmean
   }
}
```
Then we can use *mapply* function to toggle NAs in *steps*.
```{r}
data[, stepsfixed := mapply(fillna, steps, stepmean$stepmean)]
```
We can check number of NAs in *datafixed*.
```{r}
data[, sum(is.na(stepsfixed))]
```
It works.

Now we are going to re-calculate the stepsum, and plot the new histogram, with
new **mean** and **median** calculated as well.
```{r}
stepsumfixed <- data[, sum(stepsfixed, na.rm=TRUE), by=date]
setnames(stepsumfixed, c("date", "stepsumfixed"))
hist(stepsumfixed$stepsumfixed, xlab="Total Number of Steps Taken Each Day", 
     main="Histogram of the Total Number of Steps Taken Each Day(NA Fixed)",
     col="steelblue")
```

**Mean**:
```{r}
mean(stepsumfixed$stepsumfixed, na.rm=TRUE)
```

**Median**:
```{r}
median(stepsumfixed$stepsumfixed, na.rm=TRUE)
```
As we can see, after imputing missing values, the mean and median are higher 
than what were calculated before.

## Are there differences in activity patterns between weekdays and weekends?
First we need to manipulate the data a little bit.
```{r}
transformweekday <- function (datestring) {
    if (weekdays(as.Date(datestring)) %in% c("Saturday", "Sunday")) {
        factor("weekend")
    } else {
        factor("weekday")
    }
}
data[, day := sapply(date, transformweekday)]
```
The step mean can be calculated by
```{r}
stepmeancategorical <-  data[, mean(steps, na.rm=TRUE), by=c("interval","day")]
setnames(stepmeancategorical, c("interval", "day", "stepmean"))
```
Finally we get the following plot:
```{r}
library(lattice)
xyplot(stepmean ~ interval | day, 
       data=stepmeancategorical, 
       type="l", layout=c(1, 2), ylab="Number of steps")
```
