# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Read data in *.zip* format, and then convert it into a data table for program
efficiency.
```{r echo=TRUE}
library(data.table)
file <- unz("activity.zip", "activity.csv")
data <- data.table(read.table(file, header=TRUE, sep=",", na.string="NA"))
summary(data)
class(data)
```


## What is mean total number of steps taken per day?

Calculate the sum of steps group by date, and store the result into *stepsum*.
Notice that the *NA* values are removed in the *sum* function, so if the data is
missing on a specific date, it will result in a *zero sum value*.
```{r}
stepsum <- data[, sum(steps, na.rm=TRUE), by=date]
setnames(stepsum, c("date", "stepsum"))
```

We can get the histogram of the total number of steps taken each day as follows.
```{r}
hist(stepsum$stepsum, xlab="Total Number of Steps Taken Each Day", 
     main="Histogram of the Total Number of Steps Taken Each Day",
     col="blue")
```

Mean:
```{r}
mean(stepsum$stepsum, na.rm=TRUE)
```

Median:
```{r}
median(stepsum$stepsum, na.rm=TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
